---
title: "Shiny Basics"
author: "Luciano Kalabric"
date: "2024-12-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Shiny Basics
# Link https://shiny.posit.co/r/getstarted/shiny-basics/lesson1/

Installation
Shiny is an R package that makes it easy to build interactive web applications (apps) straight from R. This lesson will get you started building Shiny apps right away.

If you still haven’t installed the Shiny package, open an R session, connect to the internet, and run
Shiny comes with another package, bslib, which we’ll also use to build great looking user interfaces.

```{r, include=FALSE}
# Install libraries if not installed
list.of.packages <- c("shiny", "bslib")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

The Shiny package has eleven built-in examples that each demonstrate how Shiny works. Each example is a self-contained Shiny app.

The Hello Shiny example plots a histogram of R’s faithful dataset with a configurable number of bins. Users can change the number of bins with a slider bar, and the app will immediately respond to their input. You’ll use Hello Shiny to explore the structure of a Shiny app and to create your first app.

To run Hello Shiny directly, type:
```{r, include=FALSE}
library(shiny)
runExample("01_hello")      # a histogram
runExample("02_text")       # tables and data frames
runExample("03_reactivity") # a reactive expression
runExample("04_mpg")        # global variables
runExample("05_sliders")    # slider bars
runExample("06_tabsets")    # tabbed panels
runExample("07_widgets")    # help text and submit buttons
runExample("08_html")       # Shiny app built from HTML
runExample("09_upload")     # file upload wizard
runExample("10_download")   # file download wizard
runExample("11_timer")      # an automated timer
```

Structure of a Shiny App
Shiny apps are contained in a single script called app.R. The script app.R lives in a directory (for example, newdir/) and the app can be run with runApp("newdir").

app.R has three components:
- a user interface object
- a server function
- a call to the shinyApp function

The user interface (ui) object controls the layout and appearance of your app. The server function contains the instructions that your computer needs to build your app. Finally the shinyApp function creates Shiny app objects from an explicit UI/server pair.

# ui

Here is the ui object for the Hello Shiny example.
```{r ui, include=FALSE}
library(shiny)
library(bslib)

# Define UI for app that draws a histogram ----
ui <- page_sidebar(
  # App title ----
  title = "Hello Shiny!",
  # Sidebar panel for inputs ----
  sidebar = sidebar(
    # Input: Slider for the number of bins ----
    sliderInput(
      inputId = "bins",
      label = "Number of bins:",
      min = 1,
      max = 50,
      value = 30
    )
  ),
  # Output: Histogram ----
  plotOutput(outputId = "distPlot")
)
```
# server

Here is the server function for the Hello Shiny example.
```{r server, include=FALSE}
# Define server logic required to draw a histogram ----
server <- function(input, output) {

  # Histogram of the Old Faithful Geyser Data ----
  # with requested number of bins
  # This expression that generates a histogram is wrapped in a call
  # to renderPlot to indicate that:
  #
  # 1. It is "reactive" and therefore should be automatically
  #    re-executed when inputs (input$bins) change
  # 2. Its output type is a plot
  output$distPlot <- renderPlot({

    x    <- faithful$waiting
    bins <- seq(min(x), max(x), length.out = input$bins + 1)

    hist(x, breaks = bins, col = "#007bc2", border = "white",
         xlab = "Waiting time to next eruption (in mins)",
         main = "Histogram of waiting times")

    })

}
```

At one level, the Hello Shiny server function is very simple. The script does some calculations and then plots a histogram with the requested number of bins.

However, you’ll also notice that most of the script is wrapped in a call to renderPlot. The comment above the function explains a bit about this, but if you find it confusing, don’t worry. We’ll cover this concept in much more detail soon.

Play with the Hello Shiny app and review the source code. Try to develop a feel for how the app works. But before you do so, note that in your app.R file you will need to start with loading the Shiny package and end with a call to shinyApp:

```{r shinyapp, include=FALSE}
library(shiny)

# See above for the definitions of ui and server
ui <- ...

server <- ...

shinyApp(ui = ui, server = server)
```

Every Shiny app has the same structure: an app.R file that contains ui and server. You can create a Shiny app by making a new directory and saving an app.R file inside it. It is recommended that each app will live in its own unique directory.

You can run a Shiny app by giving the name of its directory to the function runApp. For example if your Shiny app is in a directory called my_app, run it with the following code:

```{r, include=FALSE}
setwd("C:/Users/kalab/OneDrive - FIOCRUZ/GitHub/r_analysis/redcap_analysis")
runApp("app-teste.R")
```
runGitHub
If you don’t have your own web page to host the files at, you can host your the files for free at www.github.com.

GitHub is a popular project hosting site for R developers since it does more than just host files. GitHub provides many features to support collaboration, such as issue trackers, wikis, and close integration with the git version control system. To use GitHub, you’ll need to sign up (it’s free) and choose a user name.

To share an app through GitHub, create a project repository on GitHub. Then store your app.R file in the repository, along with any supplementary files that the app uses.

Your users can launch your app by running:

```{r, include=FALSE}
runGitHub( "r_analysis", "lkalabric")
```

