---
title: "shinydashboard"
author: "Luciano K. Silva"
date: "2024-12-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## shinydashboard

Link https://rstudio.github.io/shinydashboard/get_started.html

# Installation
shinydashboard requires Shiny 0.11 or above. To install, run:

```{r installation, echo=FALSE}
# Install libraries if not installed
list.of.packages <- c("shiny", "bslib", "shinydashboard")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

# Basics

A dashboard has three parts: a header, a sidebar, and a body. Here’s the most minimal possible UI for a dashboard page.

```{r ui, echo=FALSE}
## ui.R ##
library(shinydashboard)

dashboardPage(
  dashboardHeader(),
  dashboardSidebar(),
  dashboardBody()
)
```

You can quickly view it at the R console by using the shinyApp() function. (You can also use this code as a single-file app).

```{r app, echo=FALSE}
## app.R ##
library(shiny)
library(shinydashboard)

ui <- dashboardPage(
  dashboardHeader(),
  dashboardSidebar(),
  dashboardBody()
)

server <- function(input, output) { }

shinyApp(ui, server)
```

Blank dashboard

Obviously, this dashboard isn’t very useful. We’ll need to add components that actually do something. In the body we can add boxes that have content.

```{r dashboard, echo=FALSE}
## app.R ##
library(shinydashboard)

ui <- dashboardPage(
  dashboardHeader(title = "Basic dashboard"),
  dashboardSidebar(),
  dashboardBody(
    # Boxes need to be put in a row (or column)
    fluidRow(
      box(plotOutput("plot1", height = 250)),

      box(
        title = "Controls",
        sliderInput("slider", "Number of observations:", 1, 100, 50)
      )
    )
  )
)

server <- function(input, output) {
  set.seed(122)
  histdata <- rnorm(500)

  output$plot1 <- renderPlot({
    data <- histdata[seq_len(input$slider)]
    hist(data)
  })
}

shinyApp(ui, server)
```

Basic dashboard

Next, we can add content to the sidebar. For this example we’ll add menu items that behave like tabs. These function similarly to Shiny’s tabPanels: when you click on one menu item, it shows a different set of content in the main body.

There are two parts that need to be done. First, you need to add menuItems to the sidebar, with appropriate tabNames.

```{r sidebar, echo=FALSE}
## Sidebar content
  dashboardSidebar(
    sidebarMenu(
      menuItem("Dashboard", tabName = "dashboard", icon = icon("dashboard")),
      menuItem("Widgets", tabName = "widgets", icon = icon("th"))
    )
  )
```

In the body, add tabItems with corrsponding values for tabName:

```{r body, echo=FALSE}
## Body content
  dashboardBody(
    tabItems(
      # First tab content
      tabItem(tabName = "dashboard",
        fluidRow(
          box(plotOutput("plot1", height = 250)),

          box(
            title = "Controls",
            sliderInput("slider", "Number of observations:", 1, 100, 50)
          )
        )
      ),

      # Second tab content
      tabItem(tabName = "widgets",
        h2("Widgets tab content")
      )
    )
  )
```

The default display, also shown when the “Dashboard” menu item is clicked:


